\documentclass[12pt]{article}
% or use the epsfig package if you prefer to use the old commands
\usepackage{epsfig}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{graphicx}%
\usepackage{color}
\usepackage{hyperref}
\usepackage{setspace}


%%%% UNCOMMENT NEXT BLOCK FOR CRAMMING LOTS OF TEXT ONTO PAGE
 \topskip -1cm % cram
   \oddsidemargin 5mm
   \evensidemargin 5mm
   \topmargin   -1.4cm   %1 in. from top
   \textheight 24cm % 26cm % 23cm % 22cm  (cram a bit more on the page)
   \textwidth 15.5cm
%   \parindent 0.5cm


%\newcommand{\docdb}[1]{\href{http://dayabay.ihep.ac.cn/cgi-bin/DocDB/ShowDocument?docid=#1}{\color{blue}doc-{#1}}}
\newcommand{\COtwo}{${\rm CO}_2$}

\begin{document}

\doublespacing

\title{Testing pmt calibration}

\author{David E. Jaffe \\ BNL}
\date{\today}
\maketitle
\begin{abstract}
{
Toy MC test of pmt calibration method.
}
\end{abstract}

\section{Methodology}

The ability to fit data with MC distributions of NPE (number of photoelectron) is examined.

Two different generators are used for the underlying NPE distribution. 
The poisson and betaprime functions are used. 

The betaprime function is defined as 
\begin{equation}\label{eqn:betaprime}
f(x,a,b) = \frac{x^{a-1} (1+x)^{-a-b}}{\beta(a,b)}
\end{equation}
\noindent for $x\ge 0$, $a>0$, $b>0$ with $\beta(a,b)$ defined as

\begin{equation}\label{eqn:beta}
\beta(a,b) \equiv \Gamma(a)\Gamma(b)/\Gamma(a+b)
\end{equation}
available from \href{https://docs.scipy.org/doc/scipy/reference/generated/scipy.stats.betaprime.html}{\color{blue}scipy.stats.betaprime}.

When the poisson generator is used:
\begin{itemize}
\item The MC NPE distribution is assumed to poisson with mean $\mu_M$.
\item The data NPE distributions are assumed to poisson with mean $\mu_d$ with 
	a tail fraction (tailF) of events draw from a poisson distribution with a larger mean $\mu_t$.
\end{itemize}

When the betaprime generator is used, 
\begin{itemize}
\item The MC NPE distribution is assumed to be betaprime with parameter $a = a_M$.
\item Them data NPE distribution is assumed to be betaprime with parameter $a = a_d$.
\end{itemize}
The parameter $b = 2.2$ for data and MC. The range of values chosen for $a, b$ correspond 
to distributions with mean in the range (8, 26). 
This range for $a, b$ and the betaprime generator were chosen because it resembles the NPE distribution from some of the oneton PMTs. 

The single photoelectron resolution is taken to be 0.5 PE.
The parameters and results of each configuration are shown in Table~\ref{tab:results}  for the poisson generator and 
Table~\ref{tab:betaprime_results} for the betaprime generator. (Table numbering is screwed up by latex, I don't know why). 
The best fit result is determined by a simple iterative scan of the $\chi^2$. The scanning is not optimized. 
The $\chi^2$ defined in Section~\ref{sec:chisq}.

\section{$\chi^2$ construction \label{sec:chisq}}
Consider the response of a \underline{single PMT} in the one-ton detector filled with water.
  \begin{itemize}
	\item Let $d_j = $ the sum over data events with the measured number of photoelectrons between $e_j$ and 	
	$e_{j+1}$ (in other words, the content of the $j^{th}$ bin),
	\item Let $m_j = $ the sum over MC events with the simulated number of observed photoelectrons 
	between $e_j$ and $e_{j+1}$ which is equal to the number of generated photons times the probability $a$ that
	a generated photon creates a photoelectron in the PMT.
	\item $m_j  = \sum_k^j c\times a\times g_k$ 
	where $k=$ event number, 
	$g_k = $ number of generated photons in the $k^{th}$ event and 
	$\sum_k^j$ means the sum over all entries with 
	$f\times a\times g_k$ in the $j^{th}$ bin. 
	$f=$ is the {\em calibration factor} defined such that the overall probability that a generated photon creates a photoelectron is the same as the data. 
	\item Let $M \equiv \sum_j m_j = $ the total number of MC events 
	\item and $D \equiv \sum_j d_j = $ the total number of data events for a single PMT.
	
\end{itemize}
For a single PMT, determine the {\em calibration factor} $f$ by defining the $\chi^2(f)$ as
\begin{equation}\label{eqn:chisq}
	\chi^2(f) \equiv \sum_j \bigl( \frac{d_j - m_j \frac{D}{M}}{\sigma_j} \bigr)^2
\end{equation}

$\sigma_j$ can be calculated as follows. Let $y_j = d_j - m_j\frac{D}{M}$, then

\begin{eqnarray}
\sigma_{\color{black}j}^{\color{black}2} \equiv \delta y_j^2 &= (\frac{\partial y_j}{\partial d_j} \delta d_j)^2 + (\frac{\partial y_j}{\partial m_j} \delta m_j)^2  \\
                    &=(\delta d_j)^2 + (\frac{D}{M} \delta m_j)^2 \\
                    &=(\sqrt{d_j})^2 + (\frac{D}{M} \sqrt{m_j})^2 \\
                    &= d_j + (\frac{D}{M})^2 m_j
\end{eqnarray}
Note that $\sigma_j$ depends on $f$, the calibration factor. 

Note that a sum must be taken over the number of photoelectrons in the MC events  $m_j  = \sum_k^j f\times h_k$ to evaluate $\chi^2(f)$, 
where $h_k \equiv a\times g_k$ and $\sum_k^j$ was defined above.



%\doublespacing
%\onehalfspacing   %%%% GLOBAL CHANGE OF SPACING
\clearpage
\section{Results}
Figures are provided showing the data and MC NPE distributions, best fit results, random fit results and the $\chi^2(f)$ where $f$ the calibration factor.
In general the fitted calibration factor is an unbiased estimator of the expected calibration factor for tail fractions up to 5\%. 
This conclusion applies for both the poisson and betaprime generators.
%Performance may differ if the NPE distributions are poisson or combinations of poisson distributions as approximated in this study.


\input{../FIGURES/results_table.tex}
\clearpage
\input{../FIGURES/results_betaprime_table.tex}
\clearpage

\input{ all_figint_00_30.tex}
\input{all_figint_101_105.tex}


\end{document}


